---
title: 'TCRanker preparation'
subtitle: 'Preparing input data set of TILs from PMEL mice for TCRanker'
author: "Changsheng Li"
date: '2022-05-19'
css: custom1.css
output: rmdformats::readthedown
knit: (function(inputFile, encoding) { 
      out_dir <- 'doc';
      rmarkdown::render(inputFile,
                        encoding=encoding, 
                        output_file=file.path(dirname(inputFile), out_dir, 'preparation.html')) })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

In this study, We used the paired scRNA-seq and scTCR-seq of CD8+ TILs from C57BL/6 (n=4) and PMEL (n=3) mice bearing B16-F10 melanoma tumor from the study by [Carmona et al., OncoImmunology (2020)](https://www.tandfonline.com/doi/full/10.1080/2162402X.2020.1737369)

---

#  Preparation

We'll need some packages to complete the workflow:
```{r, message=FALSE, warning=FALSE}
library(GEOquery)
library(Seurat)
library(scRepertoire)
```


And let's make the files tidy.
```{r, warning=FALSE}
setwd(getwd())
inDir <- paste0(getwd(),'/input') # Directory for input files
outDir <- paste0(getwd(),'/output') # Director for output files
dir.create(inDir)
dir.create(outDir)
```

---

# Data Set 

The corresponding GEO id of our data set is [GSE116390](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi).

```{r, cache=TRUE, warning=FALSE}
projectID <- 'Carmona.B16'
id <- 'GSE116390'
dir.create(paste0(inDir, '/', id))

supp <- getGEOSuppFiles(id, baseDir = inDir, fetch_files = F)
supp$fname
```

Among what we have in the supplementary files, what we need are the expression matrix and TCR matrix:
```{r}
supp <- supp[c(2,4),]
supp
```

---

# scRNA-seq

Firstly, we load the expression matrix:

```{r, cache=TRUE, warning=FALSE}
for(i in seq_along(supp)){
    fileDir <- paste0(inDir, '/', id, '/', supp$fname[i])
    download.file(url = supp$url[i], destfile = fileDir)
    fileFolder <- sub(".tar.gz", "", fileDir)
    untar(fileDir, exdir = fileFolder)
    rownames(supp)[i] <- fileFolder
} #Downloading and decompressing the file, might be different for other entries with different directory structures.

expMat <- Read10X(sub(".tar.*", "", rownames(supp)[2]))

query.b16 <- CreateSeuratObject(counts = expMat, project = projectID, 
                                    min.cells = 3, min.features = 50)
table(substring(colnames(query.b16), 18))
```
Table 1

Now we could assign the group name by the sample code. From the GEO description page we could find the correspondence of sample code as followed:

|sample code|1    |2  |3  |4    |5    |6  |7  |
|:----------|:----|:--|:--|:----|:----|:--|:--|
|group      |PMEL1|WT1|WT3|PMEL3|PMEL2|WT2|WT4|
<center>
Table 2
</center>
\

```{r, cache=TRUE}
libIDtoSampleID <- c('PMEL1', 'WT1', 'WT3', 'PMEL3', 'PMEL2', 'WT2', 'WT4')
names(libIDtoSampleID) <- 1:7

query.b16$Sample <- as.factor(libIDtoSampleID[substring(colnames(query.b16), 18)])

table(query.b16$Sample)
```
Table 3

Comparing three tables above, we know that the sample ID assignments were correct.

# scTCR-seq

Secondly, we load the TCR matrix:

```{r, cache=TRUE}
vdj.list <- list()
for (sampleID in 1:7) {
    sample <- libIDtoSampleID[sampleID]
    vdj <- read.csv(sprintf("%s/%s_VDJ_annotations.csv", 
                            rownames(supp)[1], sample), as.is = T)
    vdj$barcode <- paste0(sub("\\d", "", vdj$barcode), sampleID)
    
    vdj$raw_clonotype_id <- paste0(vdj$raw_clonotype_id, "-", sampleID) #Reassign sample ID to the barcodes
    vdj.list[[sampleID]] <- vdj
}

#Combine α and β chains (cdr3 segments)
combined <- combineTCR(vdj.list, ID = 1:7, sample=libIDtoSampleID, cells = "T-AB", removeNA = T, filterMulti=T)

#Strip barcodes
for (i in seq_along(combined)) {
    combined[[i]] <- stripBarcode(combined[[i]], column = 1, connector = "_", num_connects = 3)
}
```

Now we can combine the scTCR-seq data into the Seurat object
```{r}
query.b16 <- combineExpression(combined, query.b16, cloneCall = "gene", groupBy = "none")
query.b16$clonotype <- query.b16$CTaa

query.b16
colnames(query.b16@meta.data)
```

```{r}
#Save the RDS for later usage
saveRDS(query.b16, file = paste0(outDir,'/query.b16.RDS'))
```

---

Now we have the data set prepared in the format of Seurat object.

For the demonstration of how to use the TCRanker package, you can check [TCRanker demonstration](https://carmonalab.github.io/TCRanker.demo/demo.html).


