---
title: 'TCRanker demonstration'
subtitle: 'Ranking clonal exhaustion and proliferation levels of TILs from PMEL mice'
author: "Changsheng Li"
date: '2022-05-19'
css: custom2.css
output: rmdformats::readthedown
knit: (function(inputFile, encoding) { 
      out_dir <- 'doc';
      rmarkdown::render(inputFile,
                        encoding=encoding, 
                        output_file=file.path(dirname(inputFile), out_dir, 'demo.html')) })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

In this demonstration, We used the paired scRNA-seq and scTCR-seq of CD8^+^ TILs from C57BL/6 (n=4) and PMEL (n=3) mice bearing B16-F10 melanoma tumor from the study by [Carmona et al., OncoImmunology (2020)](https://www.tandfonline.com/doi/full/10.1080/2162402X.2020.1737369)

**NOTE: here we should explain in one paragraph what TCRanker does, and what we can expect from this demo.**

---

# Installation

Firstly, install TCRanker. To filter high-quality T cells, TCRanker uses [scGate](https://github.com/carmonalab/scGate), which
should also be installed.
```{r, message=FALSE, warning=FALSE}
if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
remotes::install_github("carmonalab/scGate")
remotes::install_github("carmonalab/TCRanker")

library(scGate)
library(TCRanker)
```

# Data Set

To know how to prepare the input data set for TCRanker, please refer [TCRanker preparation](https://carmonalab.github.io/TCRanker.demo/preparation.html).

```{r}
library(Seurat)

setwd(getwd()) #this command doesn't do anything ;)

outDir <- paste0(getwd(),'/output') # Director for output files
query.b16 <- readRDS(file = paste0(outDir,'/query.b16.RDS'))  ### I think we should put this file in a downloadable location, for
     #users who wants to avoid running the preprocessing script
```

---

# TCRanker

Now we could apply TCRanker over our data set.

The output data frame would have the following columns:

| columns | description
|:--------|:--------------
|clonotype|Clonotypes, the form depends on the input
|group    |Group info., optional, depends on whether in the input
|size     |Clonal size, the count of cells, also group-related (if provided)
|freq     |Clonal frequency, also group-related (if provided)
|score 1  |Clonal state score by the first gene signature 
|ranking 1|Clonal state ranking by the first gene signature 
|score 2  |Clonal state score by the second gene signature
|ranking 2|Clonal state ranking by the second gene signature 
|...      |More scores and rankings... (if more custom gene signatures provided)
<center>
Table 1
</center>
\

### Parameters: `query`， `tcr`, `filterCell`, `exhaustion` and `proliferation`

Firstly, let's see the default behaviours of TCRanker with only basic parameters
provided.

`query` is the Seurat or SingleCellExperiment object with the expression matrix
and, optimally, TCR as well.

Besides the query data set, the column names of TCR clonotype in meta.data also 
need to be specified to `tcr`. It's also possible to present the clonotype as a separated 
vector, but such practice is not recommended.

```{r, cache=TRUE}
ranking.B16 <- TCRanker(query = query.b16, tcr = "clonotype")
```

As shown above, by default, TCRanker will filter the CD8^+^ T cells first. 
It could be disabled by setting `filterCell='none'`. If you are interested,
you could also choose `filterCell='CD4T'`; `filterCell='NK'` or any other 
pre-defined models provided by [scGate](https://github.com/carmonalab/scGate)

We know that our data set contains CD8^+^ T cells only (and it's proven by how
little amount of cells filtered). So we'll set `filterCell='none'` in later
demonstration.

And this is the output data frame we'd have:

```{r, echo=FALSE}
library(DT)
DT::datatable(ranking.B16)
```
<center>
Table 2  
</center>
\

As shown above, by default, TCRanker provides the scores and rankings for 
exhaustion and proliferation. You could exclude any of them simply by setting
`exhaustion=FALSE` and `proliferation=FALSE`. Please be noted that it is 
mandatory to have at least one set of gene signature. So if you exclude both of
them, make sure to provide a custom gene signature instead.

### Parameters: `group`, `minClonSize`

The parameter `group` would lead TCRanker to add a column of group and, 
sequentially, sub-divide the entry by the group, (i.e. the cells with the same
TCR clonotype but from different groups might be divided into two or more 
sub-group and listed separately in the output.)

It's optional, yet recommended to be included in order to obtain more meaningful
result if your data is from multiple individuals or sample groups.

```{r, cache=TRUE, message=FALSE}
ranking.b16 <- TCRanker(query = query.b16, tcr = "clonotype", 
                        group = "Sample", filterCell = "none",
                        proliferation = F)


##I get this error:
# Filtering none Cells. Set filterCell=NULL to disable pre-filtering
# Error in 1:nrow(scGate.table) : argument of length 0

#filterCell should be set to NULL perhaps?

```


```{r, echo=FALSE}
DT::datatable(ranking.b16)
```
<center>
Table 3  
</center>
\

This is the output with group. Among 80 entries, there are only
`r length(unique(ranking.b16$clonotype))` unique clonotypes. You should have
noticed that the number of unique clonotypes decreased in table 3 comparing to
table 2. This is due to the `minClonSize=5` setting: when sub-divided into
smaller clonal groups, some might fall below the threshold of 5 cells to be
considered as meaningfully expanded. The number is changeable.

### Parameters: `signature`

In order to extend the functionality, custom gene signatures are also allowed
to analyse more wanted clonal states.

Here we use the signature of CD8^+^ naive-like and precursor-exhausted 
[Andreatta et al., 2021](https://www.nature.com/articles/s41467-021-23324-4)
as examples:

```{r, cache=TRUE, message=FALSE}
naiveLike <- c('Ccr7', 'Il7r', 'Sell', 'Tcf7', 'Txk', 'S1pr1', 'Lef1', 'Satb1')

Tpex <- c('Lag3', 'Xcl1', 'Crtam', 'Ifng', 'Ccl4', 'Pdcd1', 'Dusp4', 'Cd8a', 
          'Zeb2', 'Nr4a2', 'Sla', 'Nkg7', 'Tigit', 'Ctsw', 'Tnfrsf9', 'Tox',
          'Lyst', 'Tnfsf4', 'Ccl3', 'Gzmb', 'Rab27a', 'Prf1', 'Cd70', 'Plscr1',
          'Cxcl13')

customSignatures <- list(naiveLike = naiveLike, Tpex = Tpex)
```

Wrap your custom signatures in a list. And remember to name the entries to avoid
ambiguity in the output. 

Having only one set of signature, you could also use a simple vector, and the 
state would be called "user" in the output.

```{r, cache=TRUE, message=FALSE}
ranking.b16 <- TCRanker(query = query.b16, tcr = "clonotype", 
                        group = "Sample", filterCell = "none",  #NULL?
                        signature = customSignatures)
```

```{r}
#Save the RDS for later usage
saveRDS(ranking.b16, file = paste0(outDir,'/ranking.b16.RDS'))
```

```{r, echo=FALSE}
DT::datatable(ranking.b16)
```
<center>
Table 4
</center>
\

### Other parameters

- `keepObject`:
Logical, to return the `SingleCellExperiment/Saurat` object after cell filtering
or not (returned together with the ranking data frame in a list). `FALSE` by 
default.

- `assay`: 	
Name of expression data assay, By default "counts" for `SingleCellExperiment` 
and "RNA" for `Seurat`.

- `species`:
Charactor, "mouse" or "human", will be auto-detected if omitted. Currently only
these two are supported.

- `FUN`:
Function used to aggregate scores of the same clonotype. It could be mean, 
median or customized functions that take a numeric vector or list as input and 
return a single numeric. By default, it uses "mean".

- `strictFilter`:
Logical, to exclude impure cells in pure clonotype or not. Only valid when
`filterCell` is on. `TRUE` by default.

---

Now we have the clonal ranking of a certain cell state. In order to bridge the
current format of clonotypes with further lab validation or comparison, you
might need to assemble the TCR to the full sequence instead of a simple
combination of cdr3 chains.

There should be plenty of tools available. For the demonstration of how to 
generate the full sequence with `stitchr` by [Heather et al., Nucleic Acids Research (2022)](https://academic.oup.com/nar/advance-article/doi/10.1093/nar/gkac190/6553689),
you can check [TCRanker.stitchr](https://carmonalab.github.io/TCRanker.demo/stitchr.html).






